---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
activity <- read.csv("~/Documents/Coursera/Data Science/Reproducible Research/Course Project 1/activity.csv", header = TRUE, sep = ",")
activity$date <- strptime(activity$date, "%Y-%m-%d")
```


## What is mean total number of steps taken per day?

```{r}
activity2 <- data.frame("Date" = activity$date, "Steps" = activity$steps)
activity3 <- aggregate(Steps ~ Date, data = activity2, FUN = "sum")

# Plotting the histogram of total number of steps per day
with(activity3, plot(activity3$Date, activity3$Steps, main = "Total Steps per Day", xlab = "Days", ylab = "Steps", type = "h")) # "h" for type 'histogram'

mean(activity3$Steps) # result is 10766.19
median(activity3$Steps) # result is 10765
```

The mean total number of steps taken per day is 10766.19.

The median for the total number of steps taken per day is 10765.

## What is the average daily activity pattern?

# Time Series Plot for average steps per interval, averaged across all days

```{r}
activity4 <- aggregate(steps ~ interval, data = activity, FUN = "mean")
with(activity4, plot(activity4$interval, activity4$steps, main = "Time Series Plot", xlab = "5-minute Intervals", ylab = "Steps", type = "l"))
```

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
